---
title: "R and R Markdown Basics"
author: "Autumn Kloth"
date: "2024-09-06"
output: html_document
---
## Exercise 1:
In this exercise, I am creating a function with a pair of two-dimensional vectors and trying to find the Euclidean distance between the pair. 
```{r}
# mydist2D: 
  # This function takes a pair of two dimension vectors and subtracts 
  # them from each other, then squares the sum of the subtraction, summed, 
  # and then returns the square root of the whole argument to get 
  # the Euclidean distance.
# Input: 
  # I create a function called "mydist2D" with the variables a and b.
  # In the next line, I specify that a and b are vectors with specific points.
  # I also specified that when mydist2D is called, vectors a and b will return
  # by being subtracted, squared, sumed, and then square rooted. 
# Output: 
  # The output I get is 1.414214.

mydist2D <- function(a,b)
{
  return(sqrt((a[1]-b[1])^2+(a[2]-b[2])^2))
}
vec1 <- c(0,0)
vec2 <- c(1,1)
mydist2D(a=vec1,b=vec2)
```

## Exercise 2:
In this exercise, I am calculating the Euclidean distance between a pair of two vectors. I am writing a `for` loop to calculate the sum in the expression for the distance. 
```{r}
mydist <- function(a,b){
  
  l <- length(a)
  d <- 0
  
  for(i in 1:l){
    d <- d + (a[i]-b[i])^2
  }
  return(sqrt(d))
}

a <- c(0,0,0)
b <- c(1,1,1)
mydist(a,b)
```
## Exercise 3:
In this exercise, I am writing a function `evenorodd` to determine to check whether an integer is even or odd
```{r}
evenorodd <- function(x) {
  if (x %% 2 == 0) 
{
    return("even")
} else {
  return("odd")
  }
}
evenorodd(5)
```
## Exercise 4:
In this exercise, I am producing 1000 standard random variables. I am then using a function called `cumsum` that calculates the cumulative sum of the variables. I am plotting a lin of the cumulative sum in the color green. 
```{r}
x <- rnorm(1000)
y <- cumsum(x)
plot(y, type='l', col='green')
```